An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-03-19 EconGrowth-R-Functions.Rnw
Now including details of CDe parameter estimation for US and ZA.

<<cobb-douglas functions, eval=TRUE>>=
####################
# Returns an nls Cobb-Douglas model (without energy) for the country specified. 
# No energy is included in the function to be fitted.
##
cdModel <- function(countryAbbrev){
  dataTable <- loadData(countryAbbrev) #Load the data that we need.
  # Run the non-linear least squares fit to the data. No energy term desired in the Cobb-Douglas equation.
  # Establish guess values for alpha and lambda.
  lambdaGuess <- 0.0 #guessing lambda = 0 means there is no technological progress.
  alphaGuess <- 0.7 #0.7 gives good results for all countries.  
  start <- list(lambda=lambdaGuess, alpha=alphaGuess)
  # Runs a non-linear least squares fit to the data. We've replaced beta with 1-alpha for simplicity.
  model <- iGDP ~ exp(lambda*iYear) * iCapStk^alpha * iLabor^(1 - alpha)
  modelCD <- nls(formula=model, data=dataTable, start = start,
                 #Include the next 3 lines to fit with constraints.
                 #algorithm = "port",
                 #lower = list(lambda=-Inf, alpha=0),
                 #upper = list(lambda=Inf, alpha=1)
                 )
# print("str(modelCD)")
# print(str(modelCD))
# print("")
# print(modelCD$)
#   # We estimated alpha with the above code, but we have beta = 1.0 - alpha. So, get 
#   # an estimate for beta and add it to the model.
#   summaryCD <- summary(modelCD)
# print("summaryCD")
# print(summaryCD)
#   dofCD <- summaryCD$df[2] # Gives the degrees of freedom for the model.
#   tvalCD <- qt(ciHalfLevel, df = dofCD)
#   alpha <- coef(modelCD)["alpha"]
#   betaEst <- deltaMethod(modelCD, "1 - alpha") # Estimates beta and its standard error (SE).
# print("betaEst")
# print(betaEst)
#   # Calculate beta and its confidence interval and report it.
#   # Now calculate a confidence interval on beta
#   betaCICD <- with(betaEst, Estimate + c(-1.0, 1.0) * tvalCD * SE); betaCICD # Gives the confidence interval on beta.  
# print("betaCICD")
# print(betaCICD)

  return(modelCD)
}

>>>>

<<cobb-douglas functions, eval=TRUE>>=
####################
# Returns an nls Cobb-Douglas model (without energy) for the country specified. 
# No energy is included in the function to be fitted.
##
cdModel <- function(countryAbbrev){
  dataTable <- loadData(countryAbbrev) #Load the data that we need.
  # Run the non-linear least squares fit to the data. No energy term desired in the Cobb-Douglas equation.
  # Establish guess values for alpha and lambda.
  lambdaGuess <- 0.0 #guessing lambda = 0 means there is no technological progress.
  alphaGuess <- 0.7 #0.7 gives good results for all countries.  
  start <- list(lambda=lambdaGuess, alpha=alphaGuess)
  # Runs a non-linear least squares fit to the data. We've replaced beta with 1-alpha for simplicity.
  model <- iGDP ~ exp(lambda*iYear) * iCapStk^alpha * iLabor^(1 - alpha)
  modelCD <- nls(formula=model, data=dataTable, start = start,
                 #Include the next 3 lines to fit with constraints.
                 #algorithm = "port",
                 #lower = list(lambda=-Inf, alpha=0),
                 #upper = list(lambda=Inf, alpha=1)
                 )
  return(modelCD)
}

### 2013-03-19 EconGrowthSeminar.Rnw
Now including details of CDe parameter estimation for US and ZA.

\begin{frame}{Single-Factor $y$ Predictions} %----------------------------

<<Results_Table_US_SF, fig.width=presentationGraph2ColWidth, fig.height=presentationGraph1RowHeight, fig.pos="H">>=
sfUSK <- singleFactorModel("US", "K")
sfUSL <- singleFactorModel("US", "L")
sfUSQ <- singleFactorModel("US", "Q")
sfUSX <- singleFactorModel("US", "X")
sfUSU <- singleFactorModel("US", "U")
@
  \begin{columns}[c]
    \column{2.2in}
      \begin{center}
        \sfEqn
        \vspace{\baselineskip}

        \tiny
          \begin{tabular}{|c|c|c|}\hline
            $f$   & $\lambda$                      & $m$\\
            \hline
            $k$   & \Sexpr{coef(sfUSK)["lambda"]}  & \Sexpr{coef(sfUSK)["m"]}\\
            \hline
            $l$   & \Sexpr{coef(sfUSL)["lambda"]}  & \Sexpr{coef(sfUSL)["m"]}\\
            \hline
            $q$   & \Sexpr{coef(sfUSQ)["lambda"]}  & \Sexpr{coef(sfUSQ)["m"]}\\
            \hline
            $x$   & \Sexpr{coef(sfUSX)["lambda"]}  & \Sexpr{coef(sfUSX)["m"]}\\
            \hline
            $u$   & \Sexpr{coef(sfUSU)["lambda"]}  & \Sexpr{coef(sfUSU)["m"]}\\
            \hline
          \end{tabular}
        \normalsize
      \end{center}
    \column{2.2in}
<<Results_Graph_US_SF, fig.width=presentationGraph2ColWidth, fig.height=presentationGraph1RowHeight, fig.pos="H">>=
createSFLatticeGraph("US", textScaling=textScaleForGraphs, keyXLoc=keyXLoc, keyYLoc=keyYLoc)
@
  \end{columns}
  % Including at bottom of frame so as not to push other material downward on the frame.
  \putat{figure_other/flag_us.pdf}{0.055}{2.14in}{1.59in}
\end{frame}

% \begin{frame}{\includegraphics[height=0.25in]{figure_other/flag_za.png} Single-Factor Fit} %----------------------------
% <<Results_Table_ZA_SF, fig.width=presentationGraph2ColWidth, fig.height=presentationGraph1RowHeight, fig.pos="H">>=
% sfZAK <- singleFactorModel("ZA", "K")
% sfZAL <- singleFactorModel("ZA", "L")
% sfZAQ <- singleFactorModel("ZA", "Q")
% sfZAX <- singleFactorModel("ZA", "X")
% sfZAU <- singleFactorModel("ZA", "U")
% @
%   \begin{columns}[c]
%     \column{2.2in}
%       \begin{center}
%         $y = \mathrm{e}^{\lambda (t-t_0)}f^{m}$
%         \vspace{\baselineskip}
%         
%         \tiny
%           \begin{tabular}{|c|c|c|}\hline
%             $f$   & $\lambda$                      & $m$\\
%             \hline
%             $k$   & \Sexpr{coef(sfZAK)["lambda"]}  & \Sexpr{coef(sfZAK)["m"]}\\
%             \hline
%             $l$   & \Sexpr{coef(sfZAL)["lambda"]}  & \Sexpr{coef(sfZAL)["m"]}\\
%             \hline
%             $q$   & \Sexpr{coef(sfZAQ)["lambda"]}  & \Sexpr{coef(sfZAQ)["m"]}\\
%             \hline
%             $x$   & \Sexpr{coef(sfZAX)["lambda"]}  & \Sexpr{coef(sfZAX)["m"]}\\
%             \hline
%             $u$   & \Sexpr{coef(sfZAU)["lambda"]}  & \Sexpr{coef(sfZAU)["m"]}\\
%             \hline
%           \end{tabular}
%         \normalsize
%       \end{center}
%     \column{2.2in}
% <<Results_Graph_ZA_SF, fig.width=presentationGraph2ColWidth, fig.height=presentationGraph1RowHeight, fig.pos="H">>=
% createSFLatticeGraph("ZA", textScaling=textScaleForGraphs, keyXLoc=keyXLoc, keyYLoc=keyYLoc)
% @
%   \end{columns}
% \end{frame}

\begin{frame}{Cobb-Douglas} %----------------------------
  \begin{center}
    \cdEqn\\
    $\alpha + \beta = 1$\\
    $0 \leq \alpha, \beta \leq 1$
    \vspace{\baselineskip}

@@

\begin{frame}{Example Cobb-Douglas Fit} %----------------------------
  Heun complete
\end{frame}

\begin{frame}{Cobb-Douglas With Energy} %----------------------------
  \begin{center}
    \cdeEqn\\
    $\alpha + \beta + \gamma = 1$\\
    $0 \leq \alpha, \beta, \gamma \leq 1$
    \vspace{\baselineskip}

@@

\begin{frame}{Parameter Estimation for Cobb-Douglas With Energy} %----------------------------
  \begin{center}
    \cdeEqn
  \end{center}
  \begin{itemize}
    \item{Objective: determine $\lambda$, $\alpha$, $\beta$, and $\gamma$ that best fits historical data}
    \item{Almost identical to Cobb-Douglas without energy}
    \item{Perform grid search over $\gamma$}
    \item{Re-estimate $\lambda$, $\alpha$, and $\beta$ at each value of $\gamma$}
    \item{Respect sum constraint ($\alpha + \beta + \gamma = 1$)}
  \end{itemize}
\end{frame}

\begin{frame}{Grid Search Over $\gamma$}{South Africa Cobb-Douglas With Energy} %----------------------------
<<CDeGammaGraphsZA, fig.width=presentationGraph1ColWidth, fig.height=presentationGraph1RowHeight>>=
  cdeGridGraph("ZA", "Q", textScaling=0.8)
@
\end{frame}

\begin{frame}{Cobb-Douglas with Energy Example Fit} %----------------------------
  Heun complete
\end{frame}

>>>>

<<Coefs_US_SF>>=
sfUSK <- coef(singleFactorModel("US", "K"))
sfUSL <- coef(singleFactorModel("US", "L"))
sfUSQ <- coef(singleFactorModel("US", "Q"))
sfUSX <- coef(singleFactorModel("US", "X"))
sfUSU <- coef(singleFactorModel("US", "U"))
@
\begin{frame}{Single-Factor $y$ Predictions} %----------------------------
  \begin{columns}[c]
    \column{2.2in}
      \begin{center}
        \sfEqn
        \vspace{\baselineskip}

        \tiny
          \begin{tabular}{|c|c|c|}\hline
            $f$   & $\lambda$                & $m$\\
            \hline
            $k$   & \Sexpr{sfUSK["lambda"]}  & \Sexpr{sfUSK["m"]}\\
            \hline
            $l$   & \Sexpr{sfUSL["lambda"]}  & \Sexpr{sfUSL["m"]}\\
            \hline
            $q$   & \Sexpr{sfUSQ["lambda"]}  & \Sexpr{sfUSQ["m"]}\\
            \hline
            $x$   & \Sexpr{sfUSX["lambda"]}  & \Sexpr{sfUSX["m"]}\\
            \hline
            $u$   & \Sexpr{sfUSU["lambda"]}  & \Sexpr{sfUSU["m"]}\\
            \hline
          \end{tabular}
        \normalsize
      \end{center}
    \column{2.2in}
<<Results_Graph_US_SF, fig.width=presentationGraph2ColWidth, fig.height=presentationGraph1RowHeight, fig.pos="H">>=
createSFLatticeGraph("US", textScaling=textScaleForGraphs, keyXLoc=keyXLoc, keyYLoc=keyYLoc)
@
  \end{columns}
  % Including at bottom of frame so as not to push other material downward in the frame.
  \putat{figure_other/flag_us.pdf}{0.055}{2.14in}{1.59in}
\end{frame}

\begin{frame}{Cobb-Douglas (without Energy)} %----------------------------
  \begin{center}
    \cdEqn\\
    $\alpha + \beta = 1$\\
    $0 \leq \alpha, \beta \leq 1$
    \vspace{\baselineskip}

@@

\begin{frame}{Cobb-Douglas (with Energy)} %----------------------------
  \begin{center}
    \cdeEqn\\
    $\alpha + \beta + \gamma = 1$\\
    $0 \leq \alpha, \beta, \gamma \leq 1$
    \vspace{\baselineskip}

@@

\begin{frame}{Parameter Estimation for Cobb-Douglas With Energy} %----------------------------
  \begin{center}
    \cdeEqn
  \end{center}
  \begin{itemize}
    \item{Objective: determine $\lambda$, $\alpha$, $\beta$, and $\gamma$ that best fits historical data}
    \item{Almost identical to Cobb-Douglas without energy}
    \item{Perform grid search over $\gamma$}
    \item{Re-estimate $\lambda$, $\alpha$, and $\beta$ at each value of $\gamma$}
    \item{Respect sum constraint: $\alpha + \beta + \gamma = 1$}
    \item{Respect absolute constraints: $0 \leq \alpha, \beta, \gamma \leq 1$}
  \end{itemize}
\end{frame}

\begin{frame}{Grid Search Over $\gamma$}{US Cobb-Douglas With $q$} %----------------------------
<<CDeGammaGraphsUS, fig.width=presentationGraph1ColWidth, fig.height=presentationGraph1RowHeight>>=
  cdeGridGraph("US", "Q", textScaling=0.8)
@
  % Including at bottom of frame so as not to push other material downward in the frame.
  \putat{figure_other/flag_us.pdf}{0.055}{2.14in}{1.59in}
\end{frame}

<<Coefs_US_CD>>=
cdUS <- cobbDouglasData("US")
cdqUS <- cobbDouglasData("US", "Q")
cdxUS <- cobbDouglasData("US", "X")
cduUS <- cobbDouglasData("US", "U")
@
\begin{frame}{Cobb-Douglas $y$ Predictions} %----------------------------
  \begin{columns}[c]
    \column{2.2in}
      \begin{center}
        \cdeEqn
        \vspace{\baselineskip}

        \tiny
          \begin{tabular}{|c|c|c|c|c|}\hline
            $e$ & $\lambda$ & $\alpha$ & $\beta$ & $\gamma$\\
            \hline
                  & \Sexpr{cdUS["CD", "lambda"]} & \Sexpr{cdUS["CD", "alpha"]} & \Sexpr{cdUS["CD", "beta"]} & \Sexpr{cdUS["CD", "gamma"]}\\
            \hline
            $q$   & \Sexpr{cdqUS["CD", "lambda"]} & \Sexpr{cdxUS["CD", "alpha"]} & \Sexpr{cdqUS["CD", "beta"]} & \Sexpr{cdqUS["CD", "gamma"]}\\
            \hline
            $x$   & \Sexpr{cdxUS["CD", "lambda"]} & \Sexpr{cdxUS["CD", "alpha"]} & \Sexpr{cdxUS["CD", "beta"]} & \Sexpr{cdxUS["CD", "gamma"]}\\
            \hline
            $u$   & \Sexpr{cduUS["CD", "lambda"]} & \Sexpr{cduUS["CD", "alpha"]} & \Sexpr{cduUS["CD", "beta"]} & \Sexpr{cduUS["CD", "gamma"]}\\
            \hline
          \end{tabular}
        \normalsize
      \end{center}
    \column{2.2in}
<<Results_Graph_US_CD, fig.width=presentationGraph2ColWidth, fig.height=presentationGraph1RowHeight, fig.pos="H">>=
createCDLatticeGraph("US", textScaling=textScaleForGraphs, keyXLoc=keyXLoc, keyYLoc=keyYLoc)
@
  \end{columns}
  % Including at bottom of frame so as not to push other material downward in the frame.
  \putat{figure_other/flag_us.pdf}{0.055}{2.14in}{1.59in}
\end{frame}

\begin{frame}{Grid Search Over $\gamma$}{ZA Cobb-Douglas With $q$} %----------------------------
<<CDeGammaGraphsZA, fig.width=presentationGraph1ColWidth, fig.height=presentationGraph1RowHeight>>=
  cdeGridGraph("ZA", "Q", textScaling=0.8)
@
  % Including at bottom of frame so as not to push other material downward in the frame.
  \putat{figure_other/flag_za.png}{0.055}{2.14in}{1.59in}
\end{frame}

<<Coefs_ZA_CD>>=
cdZA <- cobbDouglasData("ZA")
cdqZA <- cobbDouglasData("ZA", "Q")
cdxZA <- cobbDouglasData("ZA", "X")
@
\begin{frame}{Cobb-Douglas $y$ Predictions} %----------------------------
  \begin{columns}[c]
    \column{2.2in}
      \begin{center}
        \cdeEqn
        \vspace{\baselineskip}

        \tiny
          \begin{tabular}{|c|c|c|c|c|}\hline
            $e$ & $\lambda$ & $\alpha$ & $\beta$ & $\gamma$\\
            \hline
                  & \Sexpr{cdZA["CD", "lambda"]} & \Sexpr{cdZA["CD", "alpha"]} & \Sexpr{cdZA["CD", "beta"]} & \Sexpr{cdZA["CD", "gamma"]}\\
            \hline
            $q$   & \Sexpr{cdqZA["CD", "lambda"]} & \Sexpr{cdxZA["CD", "alpha"]} & \Sexpr{cdqZA["CD", "beta"]} & \Sexpr{cdqZA["CD", "gamma"]}\\
            \hline
            $x$   & \Sexpr{cdxZA["CD", "lambda"]} & \Sexpr{cdxZA["CD", "alpha"]} & \Sexpr{cdxZA["CD", "beta"]} & \Sexpr{cdxZA["CD", "gamma"]}\\
            \hline
          \end{tabular}
        \normalsize
      \end{center}
    \column{2.2in}
<<Results_Graph_US_ZA, fig.width=presentationGraph2ColWidth, fig.height=presentationGraph1RowHeight, fig.pos="H">>=
createCDLatticeGraph("ZA", textScaling=textScaleForGraphs, keyXLoc=keyXLoc, keyYLoc=keyYLoc)
@
  \end{columns}
  % Including at bottom of frame so as not to push other material downward in the frame.
  \putat{figure_other/flag_za.png}{0.055}{2.14in}{1.59in}
\end{frame}
